Software Architectures to Integrate Workflow Engines in Science Gateways

### 2016-04-28 paper.tex
Fixed #11

\begin{abstract}
  We investigate software architectures commonly used to integrate
  workflow engines in science gateways, based on our experience with
  the CBRAIN and VIP platforms, and on our past involvement in the
  SHIWA project. We abstract software components and interactions in 6
  architectures that we use to classify 18 systems. We compare these
  architectures using metrics measuring global complexity, robustness,
  extensibility, scalability and other specific features. Results show
  the pros and cons of the different architectures: tight integration
  and sub-tasking are the simplest architectures and should therefore
  be the easiest to implement and the most robust ones, most
  architectures perform the same in terms of extensibility, the pool
  model is the most scalable and meta-workflows are only available in
  nested workflows and workflow import. Our analysis provides insights
  for software architects on how to integrate science gateways and
  workflow engines.
\end{abstract}

@@

Various science gateways have been developed, including frameworks
used as building blocks to assemble science gateways. The main
frameworks are Apache Airavata~\cite{marru2011apache}, the Catania
Science Gateway Framework~\cite{Ardizzone2012} and
WS-PGRADE/gUSE~\cite{Kacsuk2012}. Numerous science gateway instances
were built using such
frameworks~\cite{kacsuk2014science,ardizzone2012decide}. Other gateway
instances were also built independently from any framework, using
regular web development tools and libraries to access distributed
systems. CBRAIN and VIP are two such science gateways, from which we
gathered experience about the architectures described in this paper.

CBRAIN\footnote{\url{http://cbrain.mcgill.ca}}~\cite{SHER-14} is an
open-source\footnote{\url{https://github.com/aces/cbrain}}, web-based collaborative platform addressing the
challenges of data-heavy, computation-intensive neuroimaging
research. It offers transparent web-based access to remote data
sources, distributed computing sites, and an array of processing and
visualization tools, within a secure environment. CBRAIN is deployed
at a wide range of computing sites, including a network of national HPC
centers in Canada, one in Korea, one in Germany, and on numerous local
servers. As February 2016, CBRAIN has served over 420 users across 50
cities in 22 countries and has a national computing allocation of 660 CPU years
per annum.

The Virtual Imaging Platform
(VIP\footnote{\url{http://www.creatis.insa-lyon.fr/vip}})~\cite{GLAT-13}
is a web portal for medical simulation and image data analysis. It
leverages computing and storage resources available in the biomed
Virtual Organization of the European Grid Infrastructure (EGI) to
offer an open service to academic researchers worldwide. VIP
completely masks the infrastructure to enable a transparent user
experience. It uses the MOTEUR workflow engine~\cite{GLAT-13} and the
DIRAC pilot-job framework~\cite{tsaregorodtsev2008dirac} to execute
applications on EGI. Almost 900 users were registered in VIP by
April 2016 and the average CPU consumption between January 2013 and
March 2016 was 35 CPU years per month.

The architectures described hereafter also build on experience with
the SHIWA Simulation Platform~\cite{terstyanszky2014enabling}. The
SHIWA Simulation Platform is a WS-PGRADE/gUSE science gateway that was
developed in the European projects
SHIWA\footnote{\url{http://www.shiwa-workflow.eu}} and
ER-flow\footnote{\url{http://www.erflow.eu}}. It focuses on workflow
interoperability for various disciplines, in particular through the
building of meta-workflows, i.e. workflows that are built out of other
workflows. It implements the so-called Coarse-Grained Interoperability
framework where workflow engines are integrated in the science gateway
as other applications and invoked through a dedicated service. The
Fine-Grained Interoperability framework was also developed in the
SHIWA project, where conversions between workflow languages are
achieved through a common intermediate
representation~\cite{plankensteiner-prodan-etal:2013}.

>>>>

\begin{abstract}
  We investigate software architectures commonly used to integrate
  workflow engines in science gateways. We abstract software
  components and interactions in 6 architectures that we use to
  classify 18 systems. We compare these architectures using metrics
  measuring global complexity, robustness, extensibility, scalability
  and other specific features. Results show the pros and cons of the
  different architectures: tight integration and sub-tasking are the
  simplest architectures and should therefore be the easiest to
  implement and the most robust ones, most architectures perform the
  same in terms of extensibility, the pool model is the most scalable
  and meta-workflows are only available in nested workflows and
  workflow import. Our analysis provides insights for software
  architects on how to integrate science gateways and workflow
  engines.
\end{abstract}

@@

Various science gateways have been developed, including frameworks
used as building blocks to assemble science gateways. The main
frameworks are Apache Airavata~\cite{marru2011apache}, the Catania
Science Gateway Framework~\cite{Ardizzone2012} and
WS-PGRADE/gUSE~\cite{Kacsuk2012}. Numerous science gateway instances
were built using such
frameworks~\cite{kacsuk2014science,ardizzone2012decide}. Other gateway
instances were also built independently from any framework, using
regular web development tools and libraries to access distributed
systems, for instance CBRAIN~\cite{SHER-14} and VIP~\cite{GLAT-13}.
